Laves polyhedra in synthetic tennantite, Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}, and its lattice dynamics

### 2021-02-02 paper_tennantite.tex
proofreading

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 168 2019-02-25 07:15:41Z apu.v $
%%
%%
\documentclass[preprint,review,12pt]{elsarticle}

@@

\author[TISNCM]{Alexey~A.~Yaroslavzev}
\ead{yaroslavzevalex@gmail.com}
\author[MSU,ICRAS]{Alexey~N.~Kuznetsov}
\author[SIC]{Alexander~P.~Dudka}
\author[MSU]{Andrei~V.~Mironov}
\author[MIPT,TISNCM]{Sergey~G.~Buga}
\author[TISNCM]{Vladimir~V.~Denisov}

\address[TISNCM]{Technological Institute for Superhard and Novel Carbon Materials, 108840, Troitsk, Moscow, Russia}
\address[MSU]{Department of Chemistry, Lomonosov Moscow State University, 119991, Moscow, Russia}
\address[ICRAS]{Kurnakov Institute of General and Inorganic Chemistry RAS, 119991, Moscow, Russia}
\address[SIC]{Shubnikov Institute of Crystallography of Federal Scientific Research Centre “Crystallography and Photonics” of Russian Academy of Sciences, Leninskiy Prospekt 59, 119333, Moscow, Russia}
\address[MIPT]{Moscow Institute of Physics and Technology, 141700, 9 Institutsky lane, Dolgoprudny, Russia}

\begin{abstract}
Synthetic tennantite, Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}, is the analogue of an abundant mineral that belongs to the tennantite-tetrahedri\textcolor{red}{t}e group with a low lattice thermal conductivity.
\textcolor{red}{Combined data from high-quality X-ray diffraction, electron microscopy (STEM--HAADF), Raman spectroscopy, as well as DFT calculations are used to analyze the peculiarities of structure stability and dynamics.
A atomic displacement parameters and low-energy optical phonon modes are discussed within the context of the structure variations and pecularities of the charge distributions.
The latter indicates that the tennantite structure tends to conform to the covalent polar bonds and its charge distribution significantly affected by the atomic shifts in Laves polyhedra.
According to the DFT calculations, there is a population of stable model structures with varying shifts of copper atoms in Laves polyhedra, with total energies being very close (within 0.24~eV), which can explain the observed behavior of the atomic displacement parameters.}
A specially designed  technique used for the experimental analysis of the atomic displacement parameters  revealed Einstein characteristic temperatures in the tennantite structure to be in the range of 50--190~K, attributed to  low-energy optical phonon modes.

\section*{Highlights}
\begin{itemize}
  \item Synthetic tennantite Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} structure consists of disordered Laves \textcolor{red}{polyhedra}
  \item DFT calculations reveal a mixed valence of Cu atoms, and Laves \textcolor{red}{polyhedra} distortions
  \item Jahn-Teller distortions give rise to a set of unit cells with close ground energies
  \item Rattling vibration modes refer to Einstein temperatures of 74, 104, 115, 185K
\end{itemize}

@@

\begin{keyword}
tennantite Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}\sep
thermoelectic materials\sep
DFT\sep
anisotropy\sep
optical spectroscopy\sep
X-ray diffraction\sep
computer simulations\sep

@@

The tennantite-tetrahedri\textcolor{red}{t}e group represents a large class of inorganic compounds including complex copper sulfosalts with the general formula Cu\textsubscript{10+x}M\textsubscript{2-x}X\textsubscript{4}S\textsubscript{13}, where M is a transition metal (Zn, Fe, Hg, etc.), and X is a pnictogen (As, Sb, or Bi) \cite{Makovicky_2006}, with the pure tennantite itself represented by the formula Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}.
Despite being known for over two centuries, these minerals are still under investigation, both due to  structural peculiarities and potential applications. Arsenic- and antimony-containing compounds of this class are the advanced thermoelectric materials with a low lattice thermal conductivity\cite{Sootsman2009,Chetty2015}.
\textcolor{red}{For example, the studies of tetrahedrite doped with Mn, Fe, Co, Ni, Zn, Te, Se showed ZT values in the range from 0.2 to 1.13\cite{Heo2014,Suekuni2013,Lu2012,Barbier2016,Rout2020,Zhu2019}.}

\textcolor{red}{The tennantite structure treated as sphalerite-type and belongs to \hmn{I -4 3 m} space group\cite{yaroslavzev2019,Makovicky_2006}.
The structure can be described by five nonequivalent crystallographic sites: two for Cu, one for As, and two for S.
Cu1 is tetrahedrally coordinated (CuS\textsubscript{4}) by four S1 and As forms a trigonal pyramid (AsS\textsubscript{3}) with three S1.
Additionally, the S2 is centered in an octahedral unit that is named the Laves polyhedra (SCu\textsubscript{6}) with six copper atoms in sites Cu2 and Cu21.
Cu2 is trigonally coordinated (CuS\textsubscript{3}) by two S1 and one S2, Cu21 is coordinated by two S1, one S2 and one As.
Cu21 and Cu2 atoms possess very large atomic displacements perpendicular to the plane of the triangle (Fig.~\ref{fig:generel_view} right).
}
In a case of tennantite- and tetrahedrite-based materials, it is generally considered, that a low lattice thermal conductivity is caused by hybridized Cu {\it 3d}  and chalcogen \textcolor{red}{{\it p}} orbitals\cite{Lai2015}.
Such a hybridization leads to a bonding asymmetry, and out-of-plane low frequency rattling modes, which are quasi-localized and anharmonic\cite{May2016,Bouyrie2015}.

\begin{figure}[ht]
\centering
\subfigure{\includegraphics[width=0.4\columnwidth]{general_view}}
 \quad
\subfigure{\includegraphics[width=0.35\columnwidth]{general_view_laves}}
\caption{\label{fig:generel_view} \textcolor{red}{Left --- the view of the Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} tennantite structure and of coordination polyhedra around the Cu1 (blue) and the Laves polyhedra (yellow); right --- the Laves polyhedra with Cu2 and Cu21 atoms with occupancy.} }
\end{figure}

@@

Recently, we performed a multi-temperature (90--298~K) structural study of a synthetic tennantite\cite{yaroslavzev2019} and found that below ca. 120 K the copper atom distribution in tennantite changes from a more statistically delocalized Cu2 a well-defined split into Cu2 and Cu21 positions without a first-order phase transition.
This effect is accompanied by As1 and S2 atoms displacement, which was also observed in ref.~ \cite{Hathwar2019}.
We also investigated the low- and room-temperature electronic structures of a synthetic tennantite based on the DFT calculations, which confirmed the second-order transition from the low-temperature AFM to the room-temperature PM state\cite{yaroslavzev2019}.

However, the relationships between the peculiarities of the tennantite crystal and its electronic structures and thermoelectric properties, including the behavior of Laves \textcolor{red}{polyhedra} in the structure, were left aside in the previous paper\cite{yaroslavzev2019}.
In this paper, we focus on these problems, and  investigate the tennantite cell energy distribution, and the charge density differences for different Cu positions in Laves polyhedra.
Also, we consider low-energy phonon modes, and peculiarities of the atomic displacement parameters in the temperature range of 85--293~K, the same as in Ref.~\cite{Dudka2019}, in a context of the cell energy distribution.
To clarify distortions in Laves polyhedra, a study of the spherical single-crystal sample was performed and the high-angle annular dark-field (HAADF) image was obtained.

@@

High quality single crystals of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} for the XRD analysis were grown using a two-step procedure. At first, bulk Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} has been prepared by a high-temperature technique from copper (99.5~\%), arsenic (99.5~\%), and sulpher (99.5~\%).
Then, the ingots were subjected to directional recrystallization by the Bridgman-Stockbarger technique in a two-zone furnace.
The temperature in the recrystallization zone was chosen 40 K~above the tennantite melting point. A speed of the  ingot motion was 0.5~mm/hour, with the temperature gradient in the recrystallization zone about 8--10~K/mm.

\subsection{Single-crystal X-ray diffraction}\label{sec:level2}
A spherical shape single-crystal sample of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  was prepared for the anisotropic extinction study\cite{Becker1975}.
Enraf-Nonius CAD-4 diffractometer with AgK\textsubscript{$\alpha$} and graphite monochromator was used.
The spherical shape single-crystal structural study results are presented in Appendix, they are compared with the structural data \textcolor{red}{from} the \textcolor{red}{polyhedra} defined form sample\cite{yaroslavzev2019}.

\subsection{Analysis of temperature dynamics of atomic displacements}\label{sec:level2}
The temperature dynamics of an atomic displacement from the previous study\cite{yaroslavzev2019} were used to estimate Einstein characteristic temperatures by the mixed Einstein-Debye model using DebyeFit program\cite{Dudka2019}.
The data are based on the study of  Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} single-crystal sample at 85, 115, 180, 250, and 293~K.
The sample has a rectangular shape with the dimensions of 0.456~$\times$~0.31~$\times$~0.247~mm\textsuperscript{3}.
Xcalibur diffractometer equipped with a CCD EOS S2 detector (Rigaku Oxford Diffraction), MoK\textsubscript{$\alpha$} and Cobra Plus (Oxford Cryosystems) with nitrogen gas flow were used.
The original experimental methods and data processing are presented in Appendix in details.

\subsection{Theoretical calculations}\label{sec:level2}
Calculations of the stability of possible tennantite polytypes were performed based on Density Function Theory (DFT)\cite{Kohn1965} within the generalized gradient approximation (GGA) using Perdew-Burke-Ernzerhof exchange-correlation functional\cite{Perdew1996}.
We used the projector augmented wave method\cite{Blchl1994} with the periodic boundary conditions as implemented in Vienna Ab-initio Simulation Package (VASP)\cite{Kresse1996,Kresse1996-2,Kresse1993,Kresse1994}.
The plane-wave energy cut-off was set to 450~eV. To calculate the equilibrium atomic structure, the Brillouin zone was sampled according to the Monkhorst–Pack scheme\cite{Monkhorst1976} with a 6$\times$6$\times$6 grid in the k-space. The convergence criterion for the structural relaxation was set at 0.1~meV.
Both the unit cell symmetry and parameters were allowed to relax.
To facilitate unconstrained structural optimization, all the starting structures were converted to \hmn{P1} space group.

\subsection{Electron microscopy}\label{sec:level2}
High-resolution electron microscopy images were taken using FEI~Titan~80--300 transmission electron microscope in the HAADF--STEM mode (scanning translucent) with an acceleration voltage of 300~kV. The sample for microscopy was prepared using FEI~Helios~600 focused ion beam setup. HAADF image was fitted by Atomap\cite{Nord2017} package in Python.

\subsection{Raman Spectroscopy}\label{sec:level2}
Raman spectroscopy study was performed with a spectrometer TRIAX-552, equipped with Peltier TE-cooled detector CCD~SPEC~10~(Princeton Instruments).
A grating of 600~lines~mm\textsuperscript{-1}, and a 50\textsuperscript{$\times$} lens Mitutoyo~M~Plan~Apo~SL50 (a numerical aperture of 0.42) were used. Raman spectra was excited with 514.5~nm argon-ion laser Spectra-Physics Stabilite 2017 with 1~mW output power on a sample. An open argon gas flow was used to reduce air’s modes.
Calibration was performed using 520.5~cm\textsuperscript{-1} Raman peak of a polished silicon wafer. For observation of lines near the laser line we used 3~ONDAX~114-ER297-001 fiber Bragg gratings. Fittings of the Raman spectra were performed by the LMFIT\cite{LMFIT} package in Python. The fitting model was based on 6 pseudo-Voigt functions and a polynomial function for the background.

@@

\subsection{Remarks on the structure of synthetic tennantite}\label{sec:level2}
Study of the high-grade single crystal allowed us to re-investigate the crystal structure of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  with a high precision.
\textcolor{red}{The structure was refined in a way similar to described early\cite{yaroslavzev2019}.
The sum of occupancies for Cu2, Cu21 and its symmetry related site were constrained to 1.
Due to rather large distance between these atoms their ADP were refined independently.
The maximum entropy method (MEM) refinement gave the result identical to \cite{yaroslavzev2019}.}
Our data confirms that the compound contains 12 copper atoms per formula unit, and shows no evidences of a copper deficiency.
Neither XRD, nor TEM data \textcolor{red}{have shown detectable packing defects}.

@@

Due to the crystal growth method used, anisotropic extinction in the sample takes place. Extinction parameters are presented in the Table~\ref{tab:anis_ext}.
\textcolor{red}{The disorder of Cu21} is observed in the analyzed HAADF image of the sample in [110] zone (Fig.~\ref{fig:micro}).
The electron diffraction, and overlay of HAADF image with the tennantite crystal [110] zone generated by VESTA\cite{Momma2011}, are shown in Appendix.
The processed HAADF image (Fig.~\ref{fig:micro} right) shows  the ellipticity for the Cu/As atomic columns, which is a measure of the atomic column elongation\cite{Nord2017}.
The atomic column elongation is explained by the blurring of the intensity distribution, and a large atomic displacement.
A similar inhomogeneous electron density was observed in tetrahedrite\cite{Mishra2017}, and tetrahedral framework structure\cite{Suekuni2019}.
Also, the anisotropy may be caused by the crystal growth conditions or/and a crystal cell distortion.
If so, then the structure would have a diagonal distortion or/and have more or less than 12 copper atoms per cell, which is not the case.

@@

\subsection{Atomic displacement parameters analysis of the temperature dynamics based on diffraction data}\label{sec:level2}

The atomic displacement parameters (ADPs) were analyzed by DebyeFit program using the structural data.
The DebyeFit estimates the Debye or/and Einstein characteristic temperature in crystals from the equivalent atomic displacement parameters obtained at different temperatures.

At fitting, ADPs are considered as a combination of quantum zero vibrations, thermal vibrations, and static shifts\cite{Trueblood1996,Bentien2005,Nakatsuka2011}.
The thermal expansion of the crystal, and the corresponding change of the atomic vibration frequencies depend on the temperature, and may be described using a quasi-harmonic approximation\cite{willis1975thermal,Bentien2005}.
As a result ADP, is described by u\textsubscript{calc}~$=$~xu\textsubscript{E}+(1-x)u\textsubscript{D}+u\textsubscript{static}, where xu\textsubscript{E} and (1-x)u\textsubscript{D} are Einstein and Debye parts, u\textsubscript{static} is a static shift of an atom from its site in the structure model\cite{Dudka2019}.
The result of tennantite ADPs fitting is presented in the Fig.~\ref{fig:enst_temp}, and Table~\ref{tab:enst_temp}.
The average oscillation frequency of the selected atom, and its characteristic Einstein temperature, T\textsubscript{E}, can be obtained  assuming x~$=$~1.
The maximal oscillation frequency of the selected atom and its characteristic Debye temperature, T\textsubscript{D}, corresponds to the case x~$=$~0.
Distributions of zero-point oscillations, u\textsubscript{0E} or u\textsubscript{0D}, were obtained applying T~$=$~0 in the Einstein or Debye formulas.
Thus, the contribution of Einstein or Debye parts to atomic displacements consists of zero quantum vibrations, u\textsubscript{0E} or u\textsubscript{0D}, and a temperature-dependent part, u\textsubscript{E}(T) or u\textsubscript{D}(T), which can be associated with the thermal motion of atoms.
That is, u\textsubscript{E} = u\textsubscript{0E} + u\textsubscript{E}(T), and u\textsubscript{D} = u\textsubscript{0D} + u\textsubscript{D}(T). The shift of an atom from its equilibrium position due to any reason other than thermal vibrations is the sum of zero vibrations and static shift: u\textsubscript{shift} = u\textsubscript{0E} + u\textsubscript{static} or u\textsubscript{shift} = u\textsubscript{0D} + u\textsubscript{static}\cite{Dudka2019}.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.48\textwidth]{ADP_aproximation_eng}% Here is how to import EPS art
\caption{\label{fig:enst_temp} Equivalent parameters u\textsubscript{eq} of the mean-square displacements of atoms in tennantite. Points are the experimental data. The solid curves show the fit of the temperature dependences according to the expanded Einstein model.}
\end{figure}

Large static values for Cu2 atom (Table~\ref{tab:enst_temp}) \textcolor{red}{approved instability of this site\cite{Hathwar2019,yaroslavzev2019}}.
This disordering is reliably described by the anharmonic model of atomic displacements using the Gram-Charlier expansion to the 4th order.

@@

Within \textcolor{red}{approach above}, the Einstein solid model is used to describe independent harmonic oscillators.
If so, the low energy spectrum should consist of peaks with the Einstein's energies, which we found by DebyeFit (Table~\ref{tab:enst_temp}).
Results of calculations are presented in Figure~\ref{fig:low_energ_raman}.
The Einstein's energies from structural study were used as the initial points for the fit, and in Table~\ref{tab:ram_fit} the initial  and fitted wavenumbers are presented.
The low energy $l_{1}$ and $l_{2}$ peaks are well described by Einstein's energies of Cu2, Cu21, Cu1, and As (66, 40, 79, and 122 cm\textsuperscript{-1} respectively).
\textcolor{red}{Same modes were observed and analized in tetrahedrites\cite{May2016,Lai2015} with energies approximately 4, 9 and 18 meV (33, 72, 145 cm\textsuperscript{-1}).
That in good agreement with our results.
Thus, the reasons for the strong anharmonic behavior is a dynamic disorder of the copper atoms and arsenic atoms.}
Consequently, this leads to out-of-plane low frequency rattling modes, and a low lattice thermal conductivity.

>>>>

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: elsarticle-template-num.tex 168 2019-02-25 07:15:41Z apu.v $
%%
%%
\documentclass[preprint,review,12pt]{elsarticle}
\biboptions{sort&compress}

@@

\author[TISNCM]{Alexey~A.~Yaroslavzev}
\ead{yaroslavzevalex@gmail.com}
\author[MSU,ICRAS]{Alexey~N.~Kuznetsov}
\author[SIC]{Alexander~P.~Dudka}
\author[MSU]{Andrei~V.~Mironov}
\author[TISNCM]{Sergey~G.~Buga}
\author[TISNCM]{Vladimir~V.~Denisov}

\address[TISNCM]{Technological Institute for Superhard and Novel Carbon Materials, 108840, Troitsk, Moscow, Russia}
\address[MSU]{Department of Chemistry, Lomonosov Moscow State University, 119991, Moscow, Russia}
\address[ICRAS]{Kurnakov Institute of General and Inorganic Chemistry RAS, 119991, Moscow, Russia}
\address[SIC]{Shubnikov Institute of Crystallography of Federal Scientific Research Centre “Crystallography and Photonics” of Russian Academy of Sciences, Leninskiy Prospekt 59, 119333, Moscow, Russia}

\begin{abstract}
Synthetic tennantite, Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}, is the analogue of an abundant mineral that belongs to the tennantite-tetrahedri\textcolor{red}{t}e group with a low lattice thermal conductivity.
\textcolor{red}{Combined data from high-quality X-ray diffraction, electron microscopy (STEM--HAADF), Raman spectroscopy, as well as DFT calculations are used to analyze the peculiarities of its structure stability and dynamics.
Atomic displacement parameters (ADP) and low-energy optical phonon modes are discussed within the context of the structure variations and pecularities of the charge distribution.
The latter indicates that the tennantite structure tends to conform to the covalent polar bonds and its charge distribution is significantly affected by the atomic shifts in Laves polyhedra.
According to the DFT calculations, there is a population of stable model structures with varying shifts of copper atoms in Laves polyhedra, with total energies being very close (within 0.24~eV), which can explain the observed behavior of the ADP.
A specially designed  technique is used for the experimental analysis of the ADP that revealed Einstein characteristic temperatures in the tennantite structure to be in the range of 50--190~K, which is attributed to  low-energy optical phonon modes.}

\section*{Highlights}
\begin{itemize}
  \item Synthetic tennantite Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} structure consists of disordered Laves \textcolor{red}{polyhedra}
  \item DFT calculations reveal a mixed valence of Cu atoms, and Laves \textcolor{red}{polyhedra} distortions
  \item Jahn-Teller distortions give rise to a set of unit cells with close \textcolor{red}{ground} state energies
  \item Rattling vibration modes refer to Einstein temperatures of 74, 104, 115, 185K
\end{itemize}

@@

\begin{keyword}
tennantite Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}\sep
thermoelectic materials\sep
DFT calculations\sep
structural anisotropy\sep
Raman spectroscopy\sep
X-ray diffraction\sep
computer simulations\sep

@@

The tennantite-tetrahedri\textcolor{red}{t}e group represents a large class of inorganic compounds, including complex copper sulfosalts with the general formula Cu\textsubscript{10+x}M\textsubscript{2-x}X\textsubscript{4}S\textsubscript{13}, where M is a transition metal (Zn, Fe, Hg, etc.), and X is a pnictogen (As, Sb, or Bi) \cite{Makovicky_2006}, with the pure tennantite itself represented by the formula Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}.
Despite being known for over two centuries, these minerals still represent objects of intense studies, both due to  structural peculiarities and potential applications. Arsenic- and antimony-containing compounds of this class are considered as advanced thermoelectric materials with a low lattice thermal conductivity\cite{Sootsman2009,Chetty2015}.
\textcolor{red}{For instance, tetrahedrite doped with Mn, Fe, Co, Ni, Zn, Te, Se showed ZT values in the range from 0.2 to 1.13\cite{Heo2014,Suekuni2013,Lu2012,Barbier2016,Rout2020,Zhu2019}.}

\textcolor{red}{The tennantite structure is treated as a sphalerite-type, thus conforms to the \hmn{I -4 3 m} space group\cite{yaroslavzev2019,Makovicky_2006}.
The structure can be described by five nonequivalent crystallographic sites: two for Cu, one for As, and two for S.
Cu1 is tetrahedrally coordinated (CuS\textsubscript{4}) by four S1, and As forms a trigonal pyramid (AsS\textsubscript{3}) with three S1.
Additionally, the S2 is a center of an octahedral unit that is called a Laves polyhedra (SCu\textsubscript{6}) with six copper atoms in sites Cu2 and Cu21.
Cu2 is trigonally coordinated (CuS\textsubscript{3}) by two S1 and one S2, while Cu21 is coordinated by two S1, one S2 and one As atoms.
Cu21 and Cu2 atoms show very large atomic displacements perpendicular to the plane of the triangle (Fig.~\ref{fig:generel_view}, right).
}
In a case of tennantite- and tetrahedrite-based materials, it is generally considered that a low lattice thermal conductivity is caused by hybridized Cu {\it 3d}  and chalcogen \textcolor{red}{{\it p}} orbitals\cite{Lai2015}.
Such a hybridization leads to a bonding asymmetry, and out-of-plane low frequency rattling modes, which are quasi-localized and anharmonic\cite{May2016,Bouyrie2015}.

\begin{figure}[ht]
\centering
\subfigure{\includegraphics[width=0.4\columnwidth]{general_view}}
 \quad
\subfigure{\includegraphics[width=0.35\columnwidth]{general_view_laves}}
\caption{\label{fig:generel_view} \textcolor{red}{Left --- the Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} tennantite structure with coordination polyhedra around the Cu1 (blue), and the Laves polyhedra (yellow); right --- the Laves polyhedra containing Cu2 and Cu21 atoms with their partial occupancies shown as sections of spheres.} }
\end{figure}

@@

Recently, we have performed a multi-temperature (90--298~K) structural study of a synthetic tennantite\cite{yaroslavzev2019} and found that below ca. 120 K the copper atom distribution in tennantite changes from a more statistically delocalized Cu2 to a well-defined split into Cu2 and Cu21 positions without a first-order phase transition.
This effect is accompanied by As1 and S2 atoms displacement, which was also observed in \textcolor{red}{Ref.}~\cite{Hathwar2019}.
We also investigated the low- and room-temperature electronic structures of a synthetic tennantite based on the DFT calculations, which confirmed the second-order transition from the low-temperature AFM to the room-temperature PM state\cite{yaroslavzev2019}.

However, the relationships between the peculiarities of the tennantite crystal and its electronic structures and thermoelectric properties, including the behavior of Laves \textcolor{red}{polyhedra} in the structure, were put aside in the previous paper\cite{yaroslavzev2019}.
In this paper, we focus on these problems, and  investigate the tennantite cell energy distribution, and the charge density differences for different Cu positions in the Laves polyhedra.
Also, we investigate low-energy phonon modes, and peculiarities of the ADP in the temperature range of 85--293~K, the same as in Ref.~\cite{Dudka2019}, in the context of the cell energy distribution.
To clarify the distortions in Laves polyhedra, a study of the spherical single-crystal sample was performed and the high-angle annular dark-field (HAADF) images were obtained.

@@

High quality single crystals of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} for the XRD analysis were grown using a two-step procedure. At first, bulk Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} has been prepared by a high-temperature technique from copper (99.5~\%), arsenic (99.5~\%), and sulphur (99.5~\%).
Then, the ingots were subjected to directional recrystallization by the Bridgman-Stockbarger technique in a two-zone furnace.
The temperature in the recrystallization zone was chosen 40 K~above the tennantite melting point. A speed of the ingot motion was set at 0.5~mm/hour, with the temperature gradient in the recrystallization zone ca.~8--10~K/mm.

\subsection{Single-crystal X-ray diffraction}\label{sec:level2}
A spherically shaped single-crystal sample of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  was prepared for the anisotropic extinction study\cite{Becker1975}.
Enraf-Nonius CAD-4 diffractometer with AgK\textsubscript{$\alpha$} equipped with graphite monochromator was used for the data collection.
The spherical single-crystal structural study results are presented in Supplementary Material, and they are compared with the structural data \textcolor{red}{from} the \textcolor{red}{regularly shaped crystal} sample\cite{yaroslavzev2019}.

\subsection{Analysis of temperature dynamics of atomic displacements}\label{sec:level2}
The temperature dynamics of an atomic displacement from the previous study\cite{yaroslavzev2019} were used to estimate Einstein characteristic temperatures by the mixed Einstein-Debye model using DebyeFit program\cite{Dudka2019}.
The data are based on the study of  Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13} single-crystal sample at 85, 115, 180, 250, and 293~K.
The sample has a rectangular shape with the dimensions of 0.456~$\times$~0.31~$\times$~0.247~mm\textsuperscript{3}.
Xcalibur diffractometer equipped with a CCD EOS S2 detector (Rigaku Oxford Diffraction, MoK\textsubscript{$\alpha$}) and Cobra Plus (Oxford Cryosystems) apparatus with nitrogen gas flow combination were used for the data collection.
The original experimental methods and data processing are presented in Supplementary Material in details.

\subsection{Theoretical calculations}\label{sec:level2}
Calculations on the stability of possible tennantite polytypes were performed based on Density Function Theory (DFT)\cite{Kohn1965} within the generalized gradient approximation (GGA) using Perdew-Burke-Ernzerhof exchange-correlation functional\cite{Perdew1996}.
We used the projector augmented wave method\cite{Blchl1994} with the periodic boundary conditions as implemented in Vienna Ab-initio Simulation Package (VASP)\cite{Kresse1996,Kresse1996-2,Kresse1993,Kresse1994}.
The plane-wave energy cut-off was set to 450~eV. To calculate the equilibrium atomic structure, the Brillouin zone was sampled according to the Monkhorst–Pack scheme\cite{Monkhorst1976} with a 6$\times$6$\times$6 grid in the {\it k}-space. The convergence criterion for the structural relaxation was set at 0.1~meV.
Both the unit cell symmetry and parameters were allowed to relax.
To facilitate unconstrained structural optimization, all the starting structures were converted to \hmn{P1} space group.

\subsection{Electron microscopy}\label{sec:level2}
High-resolution electron microscopy images were taken using FEI~Titan~80--300 transmission electron microscope in the HAADF--STEM mode (high-angle annular dark-field scanning transmission) with an acceleration voltage of 300~kV. The sample for microscopy was prepared using FEI~Helios~600 focused ion beam setup. HAADF image was fitted by Atomap\cite{Nord2017} package written in Python.

\subsection{Raman Spectroscopy}\label{sec:level2}
Raman spectroscopy study was performed with a spectrometer TRIAX-552, equipped with Peltier TE-cooled detector CCD~SPEC~10~(Princeton Instruments).
A grating of 600~lines~mm\textsuperscript{-1}, and a 50\textsuperscript{$\times$} lens Mitutoyo~M~Plan~Apo~SL50 (a numerical aperture of 0.42) were used. Raman spectra was excited with 514.5~nm argon-ion laser Spectra-Physics Stabilite 2017 with 1~mW output power on a sample. An open argon gas flow was used to reduce air’s modes.
Calibration was performed using 520.5~cm\textsuperscript{-1} Raman peak of a polished silicon wafer. 
We used 3~ONDAX~114-ER297-001 fiber Bragg gratings to observe lines near the laser one. Fittings of the Raman spectra were performed by the LMFIT\cite{LMFIT} package written in Python. The fitting model was based on 6 pseudo-Voigt functions and a polynomial function for the background.

@@

\subsection{Remarks on the structure of synthetic tennantite}\label{sec:level2}
The study of the high quality single crystal allowed us to re-investigate the crystal structure of Cu\textsubscript{12}As\textsubscript{4}S\textsubscript{13}  with a high precision.
\textcolor{red}{The structure was refined in a way similar to previously described in Ref.~\cite{yaroslavzev2019}.
The sum of occupancies for Cu2, Cu21 and its symmetry related site were constrained to 1.
Due to rather large distance between these atoms their ADP were refined independently.
The maximum entropy method (MEM) refinement gave the result identical to \cite{yaroslavzev2019}.}
Our data confirms that the compound contains 12 copper atoms per formula unit, and shows no evidence of copper deficiency.
Neither XRD, nor TEM data \textcolor{red}{have shown detectable packing defects}.

@@

Due to the crystal growth method used, the sample shows a degree of anisotropic extinction. Extinction parameters are presented in the Table~\ref{tab:anis_ext}.
\textcolor{red}{The disorder of Cu21} is observed in the analyzed HAADF image of the sample in [110] zone (Fig.~\ref{fig:micro}).
The electron diffraction, and overlay of HAADF image with the tennantite crystal [110] zone generated by VESTA\cite{Momma2011}, are given in Supplementary Material.
Processed HAADF image (Fig.~\ref{fig:micro}, right) shows  an ellipticity for the Cu/As atomic columns, which is a measure of the atomic column elongation\cite{Nord2017}.
The atomic column elongation is evidenced by the blurring of the intensity distribution, and a large atomic displacement.
Similar effects were observed in tetrahedrite\cite{Mishra2017}, and several tetrahedral framework structures\cite{Suekuni2019}.
Also, the observed anisotropic displacement may possibly be caused by the crystal growth conditions or/and a crystal cell distortion.
However, in such case the structure must show prominent diagonal distortion and/or feature an increase or decrease of the copper content per cell (with respect to 12 copper atoms), which is not the case.

@@

\subsection{Atomic displacement parameter temperature dynamics analysis based on diffraction data}\label{sec:level2}

The ADP were analyzed by DebyeFit program using the crystallographic data.
The DebyeFit estimates the Debye or/and Einstein characteristic temperature in crystals from the equivalent ADP obtained at different temperatures.

At fitting, ADPs are considered as a combination of quantum zero vibrations, thermal vibrations, and static shifts\cite{Trueblood1996,Bentien2005,Nakatsuka2011}.
The thermal expansion of the crystal, and the corresponding change of the atomic vibration frequencies depend on the temperature, and may be described using a quasi-harmonic approximation\cite{willis1975thermal,Bentien2005}.
As a result, ADP is described by u\textsubscript{calc}~$=$~xu\textsubscript{E}+(1-x)u\textsubscript{D}+u\textsubscript{static}, where xu\textsubscript{E} and (1-x)u\textsubscript{D} are Einstein and Debye parts, u\textsubscript{static} is a static shift of an atom from its site in the structure model\cite{Dudka2019}.
The result of tennantite ADP fitting is presented in the Fig.~\ref{fig:enst_temp}, and Table~\ref{tab:enst_temp}.
The average oscillation frequency of the selected atom, and its characteristic Einstein temperature, T\textsubscript{E}, can be obtained  assuming x~$=$~1.
The maximal oscillation frequency of the selected atom and its characteristic Debye temperature, T\textsubscript{D}, corresponds to the case x~$=$~0.
Distributions of zero-point oscillations, u\textsubscript{0E} or u\textsubscript{0D}, were obtained applying T~$=$~0 in the Einstein or Debye formulas.
Thus, the contribution of Einstein or Debye parts to atomic displacements consists of zero quantum vibrations, u\textsubscript{0E} or u\textsubscript{0D}, and a temperature-dependent part, u\textsubscript{E}(T) or u\textsubscript{D}(T), which can be associated with the thermal motion of atoms.
That is, u\textsubscript{E} = u\textsubscript{0E} + u\textsubscript{E}(T), and u\textsubscript{D} = u\textsubscript{0D} + u\textsubscript{D}(T). The shift of an atom from its equilibrium position due to any reason other than thermal vibrations is the sum of zero vibrations and static shift: u\textsubscript{shift} = u\textsubscript{0E} + u\textsubscript{static} or u\textsubscript{shift} = u\textsubscript{0D} + u\textsubscript{static}\cite{Dudka2019}.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.48\textwidth]{ADP_aproximation_eng}% Here is how to import EPS art
\caption{\label{fig:enst_temp} Equivalent parameters u\textsubscript{eq} of the mean-square displacements of atoms in tennantite. Points are the experimental data. Solid curves show the fit of the temperature dependences according to the expanded Einstein model.}
\end{figure}

Large static values for Cu2 atom (Table~\ref{tab:enst_temp}) \textcolor{red}{confirmed instability of this site\cite{Hathwar2019,yaroslavzev2019}}.
This disorder is reliably described by the anharmonic model of atomic displacements using the Gram-Charlier expansion to the 4th order.

@@

Within \textcolor{red}{the approach above}, the Einstein solid model is used to describe independent harmonic oscillators.
Thus, the low energy spectrum should consist of peaks with the Einstein's energies, which we found by DebyeFit (Table~\ref{tab:enst_temp}).
The results of the calculations are presented in Figure~\ref{fig:low_energ_raman}.
The Einstein's energies from structural study were used as the initial points for the fit, with the initial and fitted wavenumbers are presented and in Table~\ref{tab:ram_fit}.
The low energy $l_{1}$ and $l_{2}$ peaks are well described by Einstein's energies of Cu2, Cu21, Cu1, and As (66, 40, 79, and 122 cm\textsuperscript{-1} respectively).
\textcolor{red}{Same modes were observed and analyzed in tetrahedrites\cite{May2016,Lai2015} with energies approximately 4, 9 and 18 meV (33, 72, 145 cm\textsuperscript{-1}) which is
in good agreement with our results.
Thus, we can state that the reason for the strong anharmonic behavior is a dynamic disorder of the copper atoms and arsenic atoms.}
Consequently, this leads to out-of-plane low frequency rattling modes, and a low lattice thermal conductivity.
